---
title: "Class 09"
author: "Sharlene Yang (PID: 16351303)"
format: pdf
---

# PDB Statistics

The main database for structural information on biomolecules. Let's see how we can use it to answer these questions:

> Download a CSV file from the PDB site (accessible from “Analyze” > “PDB Statistics” > “by Experimental Method and Molecular Type”. Move this CSV file into your RStudio project and use it to answer the following questions:

```{r}
db <- read.csv("Data Export Summary.csv")
db
```

```{r}
knitr:: kable(db)
```


> Q1: What percentage of structures in the PDB are solved by X-Ray and Electron Microscopy.


```{r}
db$X.ray
x.ray <- sum(as.numeric(gsub(",", "", db$X.ray)))
em.total <- sum(as.numeric(gsub(",", "", db$em)))

```

```{r}
#I will work with `x` as my input.

sum_comma <- function(x) {
  #substitute the comma and convert to numeric
  sum(as.numeric( gsub(",", "", x)))
}
```

For X-ray:

```{r}
sum_comma(db$X.ray) / sum_comma(db$Total)
```

For EM:

```{r}
round ( sum_comma(db$EM) / sum_comma(db$Total), 2 )
```

Q2: What proportion of structures in the PDB are protein?

```{r}
round ( sum_comma(db$Total[1]) / sum_comma(db$Total) , 2)
```


> Q3: Type HIV in the PDB website search box on the home page and determine how many HIV-1 protease structures are in the current PDB?

SKIPPED!

![HIV-PR structure from MERK with a bound drug](1HSG (1).PNG)

>Q4: Water molecules normally have 3 atoms. Why do we see just one atom per water molecule in this structure?

The structure is too low of a resolution to see the H atom. You need a sub 1 Angstrom resolution to see Hydrogen.

> Q5: There is a critical “conserved” water molecule in the binding site. Can you identify this water molecule? What residue number does this water molecule have

HOH308


# Working with Structures in R

We can use the `bio3d` package to read and perform bioinformatics calculations on PDB structures.

```{r}
library(bio3d)

pdb <-  read.pdb("1HSG")
pdb
```

```{r}
attributes(pdb)
```

```{r}
head( pdb$atom )
```

Read an ADK structure

```{r}
adk <- read.pdb("6s36")
adk
```

Perform a prediction of flexibility with a technique called NMA (normal mode analysis)

```{r}
# Perform flexiblity prediction
m <- nma(adk)
plot(m)
```

Write out a "movie" (aka trajectory) of the motion for viewing in M0lstar

```{r}
mktrj(m, file="adk_m7.pdb")
```

> Q7: How many amino acid residues are there in this pdb object?  

128 residues

> Q8: Name one of the two non-protein residues? 

HOH

> Q9: How many protein chains are in this structure

2 chains